---
title: "intermediate analysis"
author: "Sean McSweeney"
output: pdf_document
header-includes:
- \usepackage{pdflscape}
- \newcommand{\blandscape}{\begin{portrait}}
- \newcommand{\elandscape}{\end{portrait}}
geometry: legalpaper, landscape, margin=1in
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# IRI Voting Integrity Analysis

```{r message=FALSE, warning=FALSE, include=FALSE}
library(tidyverse)
library(readxl)
library(gridExtra)
library(glmnet)
library(stargazer)
library(PerformanceAnalytics)
library(corrplot)
library(Hmisc)
library(viridis)
library(RColorBrewer)
library(plm)
```

```{r message=FALSE, warning=FALSE, include=FALSE}
setwd("~/Documents/School/Grad/Spring 2022/IRI Policy Lab/IRI_Policy_Lab")
merged_datasets3 <- read_csv("merged_datasets3.csv")
```

# Exploratory Analysis

## Missing data per variable
```{r, results='asis'}
perc_complete <- bind_rows(map(merged_datasets3, ~mean(is.na(.))))
perc_complete <- perc_complete %>% 
  pivot_longer(cols = everything()) %>% 
  mutate(`%complete` = round(1 - value, digits = 2)) %>% 
  select(!value) %>% 
  arrange(-`%complete`)

perc_complete %>% write_csv("CleanedMergedData/perc_complete.csv")

<<<<<<< Updated upstream
stargazer(perc_complete, summary = FALSE, rownames = FALSE, float = FALSE, type = "text")
=======
stargazer(perc_complete, summary = FALSE, rownames = FALSE, float = FALSE)
>>>>>>> Stashed changes
```


## correlation analysis for different integrity measures:
```{r}
chart.Correlation(select(merged_datasets3, AF.el_rating, PEI.el_rating, VDEM.el_rating),
                  histogram = FALSE)
```


## Correlations between causes independent variables by category:
### Governance
```{r}
gov_indc <- c("nelda3", "nelda4", "nelda5", "POLITY5.parcomp", "FH.fh_total", "POLITY5.fragment", 
                "POLITY5.polity2", "POLITY5.durable")
c_gov <- rcorr(as.matrix(select(merged_datasets3, unlist(gov_indc))))
corrplot(c_gov$r,  method = "circle", number.cex = 0.75, type = "upper", tl.cex = 0.75, 
         order = "hclust", p.mat = c_gov$P, sig.level = c(.01, .05, 0.1),  diag = FALSE,
         insig = "label_sig", pch.cex = 1)
```



### Socioeconomic
```{r}
socioecon_indc <- c("nelda17", "nelda18", "WB.GDP_ppc", "WB.FDI", "VDEM.yrs_education",
                    "WB.inflation", "WB.gini", "WB.urban_pop", "literacy_imputed")
c_socio <- rcorr(as.matrix(select(merged_datasets3, unlist(socioecon_indc))))
corrplot(c_socio$r,  method = "circle", number.cex = 0.75, type = "upper", tl.cex = 0.75, 
         order = "hclust", p.mat = c_socio$P, sig.level = c(.01, .05, 0.1), diag= FALSE, 
         insig = "label_sig", pch.cex = 1, mar=c(0,0,2,0))
```


### Media
press_freedom

### EMB
```{r}
emb_indc <- c("PEI.EMB_cap","VDEM.EMB_cap", "nelda11", "nelda45", "nelda46")
c_emb <- rcorr(as.matrix(select(merged_datasets3, unlist(emb_indc))))
corrplot(c_emb$r,  method = "circle", number.cex = 0.75, type = "upper", tl.cex = 0.75, 
         order = "hclust", p.mat = c_emb$P, sig.level = c(.01, .05, 0.1), diag= FALSE, 
         insig = "label_sig", pch.cex = 1)
```


### pre-election violence


### ethnic divisions
```{r}
ethnic_indc <- c("VDEM.access_to_public_services_social_group", 
                 "VDEM.access_to_public_services_pol_group", 
                 "VDEM.regime_support_by_an_ethnic_group", 
                 "VDEM.is_an_ethnic_group_the_most_powerful_regime_support_group",
                 "VDEM.is_an_ethnic_group_the_most_powerful_regime_duration_group")
c_ethnic <- rcorr(as.matrix(select(merged_datasets3, unlist(ethnic_indc))))
corrplot(c_ethnic$r,  method = "circle", number.cex = 0.75, type = "upper", tl.cex = 0.3, 
         order = "hclust", p.mat = c_ethnic$P, sig.level = c(.01, .05, 0.1), diag= FALSE, 
         insig = "label_sig", pch.cex = 1, tl.srt = 45)
```


### Consequences
```{r}
conseq_indc <- c("PEI.protest", "nelda29", "VDEM.turnout", "VDEM.VAP_turnout")
c_conseq <- rcorr(as.matrix(select(merged_datasets3, unlist(conseq_indc))))
corrplot(c_conseq$r,  method = "circle", number.cex = 0.75, type = "upper", tl.cex = 0.75, 
         order = "hclust", p.mat = c_conseq$P, sig.level = c(.01, .05, 0.1), diag= FALSE, 
         insig = "label_sig", pch.cex = 1)
```




## individually regressing all variables on VDEM.el_rating
### Governance
```{r, results='asis'}
lm.nelda3 <- lm(VDEM.el_rating ~ nelda3, merged_datasets3)
lm.nelda4 <- lm(VDEM.el_rating ~ nelda4, merged_datasets3)
lm.nelda5 <- lm(VDEM.el_rating ~ nelda5, merged_datasets3)
lm.POLITY5.parcomp <- lm(VDEM.el_rating ~ POLITY5.parcomp, merged_datasets3)
lm.FH.fh_total <- lm(VDEM.el_rating ~ FH.fh_total, merged_datasets3)
lm.POLITY5.fragment <- lm(VDEM.el_rating ~ POLITY5.fragment, merged_datasets3)
lm.POLITY5.polity2 <- lm(VDEM.el_rating ~ POLITY5.polity2, merged_datasets3)
lm.POLITY5.durable <- lm(VDEM.el_rating ~ POLITY5.durable, merged_datasets3)

stargazer(title = "Governance indicators", lm.nelda3, lm.nelda4, lm.nelda5, lm.POLITY5.parcomp, 
          lm.FH.fh_total, lm.POLITY5.fragment, lm.POLITY5.polity2, 
          lm.POLITY5.durable, font.size= "small", single.row = TRUE, column.sep.width = "1pt", no.space = TRUE)
```

### Socioeconomic
```{r, results='asis'}
lm.nelda17 <- lm(VDEM.el_rating ~ nelda17, merged_datasets3)
lm.nelda18 <- lm(VDEM.el_rating ~ nelda18, merged_datasets3)
lm.WB.GDP_ppc <- lm(VDEM.el_rating ~ WB.GDP_ppc, merged_datasets3)
lm.WB.FDI <- lm(VDEM.el_rating ~ WB.FDI, merged_datasets3)
lm.VDEM.yrs_education <- lm(VDEM.el_rating ~ VDEM.yrs_education, merged_datasets3)
lm.WB.inflation <- lm(VDEM.el_rating ~ WB.inflation, merged_datasets3)
lm.WB.gini <- lm(VDEM.el_rating ~ WB.gini, merged_datasets3)
lm.WB.urban_pop <- lm(VDEM.el_rating ~ WB.urban_pop, merged_datasets3)
lm.literacy_imputed <- lm(VDEM.el_rating ~ literacy_imputed, merged_datasets3)

stargazer(title = "Socioeconomic indicators", lm.nelda17, lm.nelda18, lm.WB.GDP_ppc, lm.WB.FDI,
          lm.VDEM.yrs_education, lm.WB.inflation, lm.WB.gini,
          lm.WB.urban_pop, lm.literacy_imputed, font.size= "small", single.row = TRUE, column.sep.width = "1pt", no.space = TRUE) 

```

### Media
```{r, results='asis'}
lm.press_freedom <- lm(VDEM.el_rating ~ press_freedom, merged_datasets3)

stargazer(title = "Press indicators", lm.press_freedom)
```


### EMB
```{r, results='asis'}
lm.VDEM.EMB_cap <- lm(VDEM.el_rating ~ VDEM.EMB_cap, merged_datasets3)
lm.nelda11 <- lm(VDEM.el_rating ~ nelda11, merged_datasets3)
lm.nelda45 <- lm(VDEM.el_rating ~ nelda45, merged_datasets3)
lm.nelda46 <- lm(VDEM.el_rating ~ nelda46, merged_datasets3)

stargazer(title = "EMB indicators", lm.VDEM.EMB_cap, lm.nelda11, lm.nelda45, lm.nelda46, single.row = TRUE)
```

### ethnic divisions
```{r, results='asis'}
lm.VDEM.access_to_public_services_social_group <- lm(VDEM.el_rating ~ VDEM.access_to_public_services_social_group, merged_datasets3)
lm.VDEM.access_to_public_services_pol_group <- lm(VDEM.el_rating ~ VDEM.access_to_public_services_pol_group, merged_datasets3)
lm.VDEM.regime_support_by_an_ethnic_group <- lm(VDEM.el_rating ~ VDEM.regime_support_by_an_ethnic_group, merged_datasets3)
lm.VDEM.is_an_ethnic_group_the_most_powerful_regime_support_group <- lm(VDEM.el_rating ~ VDEM.is_an_ethnic_group_the_most_powerful_regime_support_group, merged_datasets3)
lm.VDEM.is_an_ethnic_group_the_most_powerful_regime_duration_group <- lm(VDEM.el_rating ~ VDEM.is_an_ethnic_group_the_most_powerful_regime_duration_group, merged_datasets3)

stargazer(title = "Ethnic indicators", lm.VDEM.access_to_public_services_social_group, lm.VDEM.access_to_public_services_pol_group, lm.VDEM.regime_support_by_an_ethnic_group, 
          lm.VDEM.is_an_ethnic_group_the_most_powerful_regime_support_group, lm.VDEM.is_an_ethnic_group_the_most_powerful_regime_duration_group,
          single.row = TRUE)
```

### Consequences
```{r, results='asis'}
lm.PEI.protest <- lm(VDEM.el_rating ~ PEI.protest, merged_datasets3)
lm.nelda29 <- lm(VDEM.el_rating ~ nelda29, merged_datasets3)
lm.VDEM.turnout <- lm(VDEM.el_rating ~ VDEM.turnout, merged_datasets3)
lm.VDEM.VAP_turnout <- lm(VDEM.el_rating ~ VDEM.VAP_turnout, merged_datasets3)

stargazer(title = "Consequnces indicators", lm.PEI.protest, lm.VDEM.turnout, lm.VDEM.VAP_turnout)
```

\newpage

# Trying models! 

## Subgrouping income/democracy
Dividing the country-year's into three different development levels by income per capita. Also dividing country-year's into different levels of democracy/autocracy.  
Income per capita levels:  
4: (3140, 16400]  
3: (1220, 3140]  
2: (627, 1220]  
1: [262, 627]  

FreedomHouse levels:  
4: (11, 14]  
3: (9, 11]  
2: (7, 9]  
1: [2, 7]  

```{r echo=FALSE, message=FALSE, warning=FALSE}
ggplot(merged_datasets3) + geom_histogram(aes(WB.GDP_ppc))
ggplot(merged_datasets3) + geom_histogram(aes(FH.fh_total))
ggplot(merged_datasets3) + geom_point(aes(WB.GDP_ppc, FH.fh_total))


merged_datasets3 <- merged_datasets3 %>% 
  mutate(income_level = cut_number(WB.GDP_ppc, n = 4, labels = FALSE),
         fh_level = cut_number(FH.fh_total, n = 4, labels = FALSE)) %>% 
  mutate(income_level = case_when(income_level == 4 ~ "$16,400 - $3,140",
                                  income_level == 3 ~ "$3,140 - $1,220",
                                  income_level == 2 ~ "$1,220 - $627",
                                  income_level == 1 ~ "$627 - $262"),
         fh_level = as.factor(fh_level))

merged_datasets3$income_level <- factor(merged_datasets3$income_level,
                         levels = c("$16,400 - $3,140", "$3,140 - $1,220", "$1,220 - $627", "$627 - $262"))

```


## EMB capacity models
```{r warning=FALSE, results='asis'}
ggplot(merged_datasets3) + 
  geom_point(aes(VDEM.EMB_cap, VDEM.el_rating)) + 
  theme_minimal() + 
  labs(x = "Election Management Body capacity", 
       y = "Election integrity perception",
       title = "The relationship between free and fair elections, and EMB capacity")

stargazer(lm(VDEM.el_rating ~ VDEM.EMB_cap + income_level, merged_datasets3))
ggplot(drop_na(merged_datasets3, income_level), aes(VDEM.EMB_cap, VDEM.el_rating, color = income_level)) + 
  geom_point() + 
  theme_minimal() + 
  labs(x = "Election Management Body capacity", 
       y = "Election integrity perception",
       title = "The relationship between free and fair elections, and EMB capacity") + 
  geom_smooth(method = "lm", se = FALSE)

stargazer(lm(VDEM.el_rating ~ VDEM.EMB_cap + fh_level, merged_datasets3))
ggplot(drop_na(merged_datasets3, fh_level), aes(VDEM.EMB_cap, VDEM.el_rating, color = fh_level)) + 
  geom_point() + 
  theme_minimal() + 
  labs(x = "Election Management Body capacity", 
       y = "Election integrity perception",
       title = "The relationship between free and fair elections, and EMB capacity") + 
  geom_smooth(method = "lm", se = FALSE)

```

## lasso modeling
```{r eval=FALSE, include=FALSE}
# lasso analysis
lm(VDEM.elction_free_fair ~ VDEM.EMB_capacity_osp, merged_datasets3)
lm(VDEM.elction_free_fair ~ `PEI.Electoral authorities index (0-100), imputed`, merged_datasets3)
lm(VDEM.elction_free_fair ~ VDEM.EMB_capacity_osp, merged_datasets3)

sapply(dt_split, function(x) sum(is.na(x)))

merged_lasso <- merged_datasets3 %>% 
  select(!c(country_year:`PEI.Rating of electoral integrity (1-10)`, 
            `PEI.Voting results/reactions index (protests/disputes) (0-100), imputed`,
            `PEI.Electoral authorities index (0-100), imputed`,
            `PEI.turnout`, `WB.Gini index`, `POLITY5.polity2`,
            nelda46, nelda30,POLITY5.fragment, POLITY5.durable, POLITY5.parcomp, `WB.Inflation, consumer prices (annual %)`,
            VDEM.yrs_education, `Press Freedom Index`)) %>%
  drop_na() 


dt_split <- merged_lasso %>%
  mutate(random = rnorm(nrow(merged_lasso))) %>%
  arrange(random) %>%
  mutate(set = case_when(
    row_number() <= nrow(merged_lasso)*0.75 ~ "training",
    row_number() > nrow(merged_lasso)*0.25 &  row_number() <= nrow(merged_lasso)*1 ~ "test")
  ) %>% 
  mutate(Y = VDEM.elction_free_fair) %>%
  dplyr::select(!VDEM.elction_free_fair) %>% 
  relocate(Y, .before = VDEM.EMB_capacity_osp)

train <- dt_split %>%
  filter(set == "training") %>%
  dplyr::select(-c(random, set))

test <- dt_split %>%
  filter(set == "test") %>%
  dplyr::select(-c(random, set))

x = as.matrix(train[,2:25]) 
y = as.matrix(train[,1])

lasso.cv = cv.glmnet(x, y)
coef.min = coef(lasso.cv, s=lasso.cv$lambda.min)

plot(lasso.cv)

plot(lasso.cv$glmnet.fit , xvar = "lambda")
abline(v = log(lasso.cv$lambda.min), lty = 2, col = "red")
abline(v = log(lasso.cv$lambda.1se), lty = 2, col = "green")
legend("topright", legend = c("min", "1se"), lty = 2, col = c("red", 
                                                              "green"))

coef(lasso.cv, s = c(lasso.cv$lambda.min))
as_tibble(coef(lasso.cv, s = c(lasso.cv$lambda.min)), validate = NULL, .name_repair = NULL)

# capture with Lasso only the ones that min the lamdba
coefList <- coef(lasso.cv, s='lambda.min')
coefList <- data.frame(coefList@Dimnames[[1]][coefList@i+1],coefList@x)
coefList <- coefList[2:19,]

# capture with Lasso only the ones that 1se the lamdba
coefList2 <- coef(lasso.cv, s='lambda.1se')
coefList2 <- data.frame(coefList2@Dimnames[[1]][coefList2@i+1],coefList2@x)
coefList2 <- coefList2[2:9,]

# lasso model that min lambda
lm.lasso.min <- lm(VDEM.el_rating ~ VDEM.EMB_cap + VDEM.turnout + nelda17 + nelda3 + nelda4 + nelda5 + nelda29 + 
                     FH.fh_total + WB.FDI + WB.urban_pop +  literacy_imputed + VDEM.elec_viol + VDEM.international_monitors_present + 
                     VDEM.access_to_public_services_pol_group +  VDEM.regime_support_by_an_ethnic_group + 
                     VDEM.is_an_ethnic_group_the_most_powerful_regime_duration_group + 
                     VDEM.is_an_ethnic_group_the_most_powerful_regime_support_group, data = merged_datasets3)


# lasso model that use 1se lambda
lm.lasso.1se <- lm(VDEM.el_rating ~ VDEM.EMB_cap + nelda3 + nelda29 + FH.fh_total + WB.urban_pop + 
                     `VDEM.elec_viol` + VDEM.international_monitors_present + VDEM.access_to_public_services_pol_group, data = merged_datasets3)

stargazer(lm.lasso.1se, lm.lasso.min, title = "lasso models", font.size= "small", column.labels =c("min 1se", "min lambda"),
          column.sep.width = "1pt", no.space = TRUE, omit.stat=c("LL","ser"), align = TRUE, single.row =TRUE)
```

## fixed effects
create panel dataset
```{r, results='asis'}
merged.p <- pdata.frame(select(merged_datasets3, !country_year), index = c("country", "year"))
fixedeff <- plm(VDEM.el_rating ~ VDEM.EMB_cap, data = merged.p, model = "within")
stargazer(fixedeff, lm(VDEM.el_rating ~ VDEM.EMB_cap, merged_datasets3))
```

Should make a linechart with year on x and el_rating on y. With each country as a line. Or maybe just IRI countries as a line. This might motivate a panel data chart.
```{r warning=FALSE}
iri_countries <- c("Kenya", "Chad", "Somalia", "Sudan", "Nigeria", 
                   "Sierra Leone", "Congo", "Zimbabwe", "Angola", "Libya", 
                   "Congo Kinshasa", "Congo (Kinshasa)", "Congo, Dem. Rep.")

ggplot(filter(merged_datasets3, country %in% iri_countries)) + 
  geom_line(aes(year, VDEM.el_rating, color = country))

ggplot(merged_datasets3) + 
  geom_line(aes(year, VDEM.el_rating, color = country))

#facet by income level and FH level
```


\newpage
## Lagged dependent variable

```{r, results='asis'}
lagged.lm <- lm(VDEM.el_rating ~ VDEM.el_rating.lagged, merged_datasets3)
#stargazer(lagged.lm)

lagged.lm.incomelevel <- lm(VDEM.el_rating ~ VDEM.el_rating.lagged + income_level, merged_datasets3)
#stargazer(lagged.lm.incomelevel)

lagged.lm.income <- lm(VDEM.el_rating ~ VDEM.el_rating.lagged + WB.GDP_ppc, merged_datasets3)
lagged.lm.fh <- lm(VDEM.el_rating ~ VDEM.el_rating.lagged + FH.fh_total, merged_datasets3)
lagged.lm.fhlevel <- lm(VDEM.el_rating ~ VDEM.el_rating.lagged + fh_level, merged_datasets3)

stargazer(lagged.lm, lagged.lm.incomelevel, lagged.lm.income,lagged.lm.fh,lagged.lm.fhlevel)


p.income <- ggplot(merged_datasets3, 
       aes(VDEM.el_rating.lagged, VDEM.el_rating, color = log(WB.GDP_ppc))) + 
  geom_point() + 
  theme_minimal() + 
  labs(x = "previous election's integrity rating (lagged)", 
       y = "election integrity rating") + 
   #    title = "To predict the future, look to the past") + 
    scale_color_continuous(name = "GDP per capita (log)", type = "viridis", labels=scales::dollar_format())


ggplot(merged_datasets3, 
       aes(VDEM.el_rating.lagged, VDEM.el_rating, color = income_level)) + 
  geom_point() + 
  theme_minimal() + 
  labs(x = "previous election's integrity rating (lagged)", 
       y = "election integrity rating", 
       title = "To predict the future, look to the past")


p.fh <- ggplot(merged_datasets3, 
       aes(VDEM.el_rating.lagged, VDEM.el_rating, color = FH.fh_total)) + 
  geom_point() + 
  theme_minimal() + 
  labs(x = "previous election's integrity rating (lagged)", 
       y = "election integrity rating") +  
      # title = "To predict the future, look to the past") + 
  scale_color_continuous(name = "Freedom House Scores \n(lower score = more free)", type = "viridis")

grid.arrange(p.income, p.fh, ncol = 2)
```




